---
title: Voter Registration by Counties
author: Jake and Katie
---

```{r}
library(here)
library(tidyverse)
library(usmap)
library(maptools)
library(readxl)

library(tigris)
library(tmap)
library(tmaptools)

library(maps)
library(mapdata)
```


```{r}
load(file=here::here("Analysis","vregdat1.rda"),verbose=TRUE)
vregdat1$fips <- vregdat1$FIPSCode
vregdat1$state <- vregdat1$State_Abbr

```

What counties have the lowest voter registration rates overall? (Here are the
lowest 100 out of `r nrow(vregdat1)`.

```{r}

vregdat1  %>% dplyr::select(geoname,State_Abbr,tot_est_Total,same_day_state,automatic_state,vratereg,cvap_est_Total,totreg,totact) %>% arrange(vratereg) %>% print(.,n=100)

```

What are the 100 lowest registration counties among stats with neither same day
nor automatic registration?



```{r}

vregdat1  %>% filter(same_day_state==0 & automatic_state==0) %>% dplyr::select(geoname,State_Abbr,tot_est_Total,vratereg,cvap_est_Total,totreg,totact) %>% arrange(vratereg) %>% print(.,n=100)

```

Is there variation within Pennsylvania and/or Utah in registration by County?

```{r}

vregdat1  %>% filter(State_Abbr %in% c("PA","UT")) %>%
 dplyr::select(geoname,State_Abbr,tot_est_Total,vratereg,cvap_est_Total,totreg,totact) %>%
 arrange(State_Abbr,vratereg) %>% print(.,n=100)

vregdat1  %>% filter(State_Abbr %in% c("PA","UT")) %>% group_by(State_Abbr) %>% summarize(min(vratereg),quantile(vratereg,.25),median(vratereg),quantile(vratereg,.75),max(vratereg))


bp_dat <- vregdat1 %>% filter(!is.na(vratereg) & vratereg < 2) %>% mutate(mean_reg = mean(vratereg,na.rm=TRUE)) %>% arrange(State_Abbr,desc(mean_reg))

```

Boxplot:

```{r}

vregdat1$same_day_stateF <- factor(vregdat1$same_day_state,labels=c("No Same Day Reg","Same Day Reg"))
vregdat1$automatic_stateF <- factor(vregdat1$automatic_state,labels=c("No Automatic Reg","Automatic Reg"))

with(vregdat1,table(same_day_state,same_day_stateF,exclude=c()))
with(vregdat1,table(same_day_state,automatic_stateF,exclude=c()))

bplot_dat <- vregdat1 %>% group_by(State_Abbr) %>% filter(!is.na(vratereg) & vratereg < 2 & State_Abbr!="DC") %>% summarize(min=min(vratereg),q25=quantile(vratereg,.25),med=median(vratereg),mean=mean(vratereg),q75=quantile(vratereg,.75),max=max(vratereg),easy_reg=unique(interaction(same_day_stateF,automatic_stateF))) %>% arrange(desc(med)) %>% print(n=100)
bplot_dat$stateF <- factor(bplot_dat$State_Abbr)

bp <-  ggplot(bplot_dat) +
    geom_boxplot(aes(y=reorder(stateF,mean),
                 xmax = max,
                 xupper = q75,
                 x = med,
                 xmiddle = med,
                 xlower = q25,
                 xmin = min), stat = "identity") +
         theme_bw() +
         facet_wrap(~easy_reg,scales="free")+
         xlim(.1,1.2) +
         ylab("") +
         xlab("Proportion of Citizen of Voting Age Registered to Vote (EAC+Census Data)")
bp

ggsave(bp,file=here::here("Analysis","state_reg_boxplot.png"),dpi=600)

```

Think about some maps for slides on Wednesday. Ben says to use the tigris
approach.

```{r}

map_dat <- vregdat1 %>% dplyr::select(fips,vratereg,same_day_state,automatic_state,state,geoname) %>% filter(vratereg < 2)
map_dat$vratereg_rank <- rank(map_dat$vratereg)

full_map_rank <- plot_usmap(regions="counties",data=map_dat,values="vratereg_rank")

full_map_rank

not_same_not_auto_map_rank <- plot_usmap(regions="counties",data=filter(map_dat,same_day_state==0&automatic_state==0),values="vratereg_rank")
not_same_not_auto_map_rank


not_same_not_auto_map <- plot_usmap(regions="counties",data=filter(map_dat,same_day_state==0&automatic_state==0),values="vratereg")
not_same_not_auto_map

not_same_not_auto_map

small_map_dat <- filter(map_dat,state %in% c("PA","MI", "VA", "WA")) 

small_map_dat <- small_map_dat %>%  mutate(vratereg_log=log(vratereg),
        vratereg_exp=exp(vratereg))

small_map <- plot_usmap(regions="counties",
    data=small_map_dat,
    include=c("PA","MI","VA","WA"),
    values="vratereg")+
scale_fill_continuous(type="viridis", name = "% Reg (2020)"#, label = scales::comma
)

small_map_2 <- small_map + coord_cartesian() + facet_grid(cols=vars(state), drop = T, scales="free", space="free")
small_map_2

ggsave(small_map_2,file=here::here("Analysis","selected_states_counties.png"),dpi=600)

small_counties <- map_data('county') %>% filter(region %in% c("pennsylvania","utah"))

```




```{r}
bplot_dat$selected_states <- factor(as.numeric(bplot_dat$stateF %in% c("PA","MI", "VA", "WA")))


bp_small <-  ggplot(bplot_dat,aes(color=selected_states,fill=selected_states)) +
    geom_boxplot(aes(y=reorder(stateF,mean),
                 xmax = max,
                 xupper = q75,
                 x = med,
                 xmiddle = med,
                 xlower = q25,
                 xmin = min), stat = "identity") +
         theme_bw() +
         facet_wrap(~easy_reg,scales="free")+
         xlim(.1,1.2) +
         ylab("") +
         xlab("Proportion of Citizen of Voting Age Registered to Vote (EAC+Census Data)") + theme(legend.position = "none")
bp_small

ggsave(bp_small,file=here::here("Analysis","selected_state_reg_boxplot.png"),dpi=600)



```

